---
title: "OilGas"
output: html_document
---


## Libraries 

```{r}
library(tidyverse)
library(rstatix)
library(leaps)
library(janitor)
library(dplyr)
library(stringr)
```


## Import the Dataset 

```{r}
SWIW_data <- read_csv("../data/Oil&GasWells.csv")
SWIW_df <- SWIW_data[sample(nrow(SWIW_data), 0.10 * nrow(SWIW_data)), ]
```


## Data Scrubbing and Transformation (Cleaning)

```{r}
SWIW_df_clean <- SWIW_df %>% mutate(across(ends_with("Date") | starts_with("Well_Date"), as.Date, format = "%m/%d/%Y")) %>% mutate(across(c(County, `Well Type`, `Well Status`), ~trimws(toupper(.)))) %>%
mutate(across(c(County, `Well Type`, `Well Status`), as.factor)) %>% mutate(year_approved = lubridate::year(Well_Date_Approved), years_active = as.numeric(difftime(Sys.Date(), Well_Date_Complete, units = "days")) / 365)

SWIW_df_clean <- janitor::clean_names(SWIW_df_clean)
DT::datatable(SWIW_df_clean)
```

```{r}
# normalized well_id for join across ODNR files
normalize_well_id <- function(x) {
  x %>%
    as.character() %>%
    str_trim() %>%
    str_to_upper() %>%
    str_replace_all("[^A-Z0-9]", "")
}

SWIW_df_clean <- SWIW_df_clean %>%
  mutate(
    well_id = normalize_well_id(permit_number_api)
  )


# Handle missing fields for key numeric variables
key_numeric <- c("elevation", "td_md", "du_acres")

existing_numeric <- intersect(names(SWIW_df_clean), key_numeric)

SWIW_df_clean <- SWIW_df_clean %>%
  group_by(county) %>%
  mutate(
    across(
      all_of(existing_numeric),
      ~ ifelse(is.na(.), median(., na.rm = TRUE), .)
    )
  ) %>%
  ungroup()

# ODNR units (feet to meters), same original columns
SWIW_df_clean <- SWIW_df_clean %>%
  mutate(
    elevation_m = if ("elevation" %in% names(.)) elevation * 0.3048 else NA_real_,
    td_md_m     = if ("td_md" %in% names(.))     td_md     * 0.3048 else NA_real_
  )
```


## EDA 

```{r}
head(SWIW_df_clean)
```

```{r}
library(maps)

data_points <- data.frame(
  lon = SWIW_df$`Well Longitude`,  
  lat = SWIW_df$`Well Latitude`,
  county = SWIW_df$County 
)

# Get map data for Ohio counties
ohio_counties <- map_data("county") %>%
  filter(region == "ohio")

ggplot() +
  geom_polygon(
    data = ohio_counties,
    aes(x = long, y = lat, group = group),
    fill = "gray95", color = "black", linewidth = 0.3
  ) +
  geom_point(
    data = data_points,
    aes(x = lon, y = lat),
    color = "#0072B2", alpha = 0.7, size = 2
  ) +
  coord_fixed(1.3) +
  theme_minimal(base_size = 12) +
  labs(
    title = "Coordinates by County in Ohio",
    x = "Longitude", y = "Latitude"
  ) +
  theme(
    legend.position = "none",        # remove legend
    panel.grid = element_blank(),    # remove gridlines
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    axis.text = element_text(color = "black")
  )
```

The following map gives the well locations across Ohio counties in term of longitude, latitude coordinates. The way that this can be used is to compare to other maps that give insight into the disposal issue. For example, a map of wells that go out of commission can be overlaid to see which regions are affected the most by flooding. 

```{r}

# Numerical Summary Statistics 

Well_Stats_Count <-  SWIW_df_clean %>% count(well_status, sort = TRUE) %>%
  mutate(Percent = round(100 * n / sum(n), 2))
print(Well_Stats_Count)

## Note: Colorcode wells onto map to triganulate which wells are "plugged and abandonded" in Ohio. 

SWIW_df_clean %>%
  group_by(well_status) %>%
  get_summary_stats(
    vars = c("elevation", "du_acres", "well_ip_gas", "well_ip_oil", "well_total_depth"),
    type = "full"
  )

SWIW_df_clean %>%
  get_summary_stats(
    vars = c("elevation", "du_acres", "well_ip_gas", "well_ip_oil", "well_total_depth"),
    type = "full"
  )
```

